---
title: Some basic approaches to analyze a simple simulated data example
author: 'David Bock'
date: '2020-11-16'
slug: post-2020-11-16
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2020-11-16T07:09:24+01:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---





```{r echo = FALSE , include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE , message = FALSE)

library(tidyverse)
library(broom)
library(knitr)
library(brms)

```


## Introduction

In this post I will explore some basic approaches to analyze a simple simulated data example. 

## Two sample continuous outcome
A depend variable Y and a indedependent factor X (two levels, A and B). 
I standardize Y to get zero mean and unit standard deviation. 

```{r A,  include=T }
DAT <- tibble(N = 60 , 
              X = rbinom(N, 1, 0.5),
              Y =  0.5*X + rnorm(N)) %>% 
        mutate(X = factor(X, labels = c("A", "B")) , 
               Y = (Y-mean(Y, na.rm=T))/sd(Y, na.rm=T))
```

Describe the data (recall that Y has been standardized)

```{r B,  include=T}
DESC <- DAT %>%   group_by(X) %>%
  summarise( mean = mean(Y, na.rm=T) ,
             sd = sd(Y, na.rm=T) ,
             min=min(Y, na.rm=T),
             max=max(Y, na.rm=T) ,
             n = n()) %>%
  kable(digits = 2)
DESC
```

Plot the data
```{r C,  include=T}

PLT_1 <- DAT %>%          ggplot(  aes(x=Y, color=X , fill=X )) +
  #   geom_histogram(aes(y=..density..), position="identity", alpha=0.2 , binwidth= 0.333)+
  geom_density(alpha=0.2)+
  scale_color_manual(values=c("red", "blue"))+
  scale_fill_manual(values=c("red", "blue"))+
  labs(x="Y", y = "Density")+
  theme_classic() +
  geom_rug() +
  theme(legend.text = element_text(size = 12, colour = "black", angle = 0)) +
  theme(legend.position = c(0.8, 0.8)) +
  theme(axis.text.x = element_text(size=12), axis.text.y = element_text(size=12))


```


```{r echo = FALSE , include=T , comment=NA} 
PLT_1
```



## Frequentist inference by a standard linear model (two-sample t-test):
```{r D,  include=T}
m1 <- tidy(lm(Y~X , data = DAT), conf.int = T,  conf.level = 0.95)[2,]
M <- signif(as.numeric(m1[, 2]) , digits = 2) ; 
L<- signif(as.numeric(m1[, 6]) , digits = 2) ; 
U <- signif(as.numeric(m1[, 7]), digits = 2) ; 
U_ <- paste( U,"(95%CI:",L, ";",U, ")")

```
Mean difference and 95% compatibility intervals:

```{r echo = FALSE , include=T , comment=NA} 
U_
```

